### 2018-10-13 4_prepare_manuscript/tex/manuscript.tex

%Notes to self on compiling without BibTeX
%Compile original file as normal with BibTex, using bibunits
%copy contents of bu1.bbl and paste into bibliography section
%comment out \usepackage{bibunits} in preamble
%comment out begin and end bibunit, and putbib
%compile twice with XeLateX

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode
\documentclass[reqno,12pt]{amsart}
\usepackage[foot]{amsaddr}
\usepackage{graphicx}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage[paperwidth=8.5in,paperheight=11in,text={6in,9in},left=1.25in,top=1in,headheight=0.25in,headsep=0.4in,footskip=0.4in]{geometry}

%citation format
\usepackage[authoryear]{natbib}
%use authoryear in brackets for conventional format
% use "super" in brackets for numbers to biblio

% \bibpunct[, ]{}{}{,}{s}{}{,}
\bibpunct[, ]{(}{)}{,}{a}{}{,}
%change s to a, and put open and closed parentheses in first two bracketed arguments for author-year citations

\usepackage{subfigure}
\usepackage{lineno}
\usepackage{threeparttable}

\usepackage{comment} %make block comments

\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}
\DeclareMathOperator{\E}{E}

%get rid of red link boxes
\usepackage{hyperref}
%\usepackage{endfloat}
\hypersetup{
    colorlinks,
    linkcolor={red!80!black},
    citecolor={blue!90!black},
    urlcolor={blue!90!black}
}

%flexible hyperrefs
%\usepackage{cleveref}

%gives ability to fix figure locations with H
\usepackage{float}

%allows subsetting of bibliography units
%\usepackage{bibunits}

\synctex=1

\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}% 
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

%\usepackage{lmodern}
%\usepackage{unicode-math}
\usepackage{amsmath}
\usepackage{mathspec}
\usepackage{units}
\usepackage{xltxtra}
\usepackage{xunicode}
\defaultfontfeatures{Mapping=tex-text}
%\setmainfont[Ligatures={Common}]{EB Garamond}
%\setmathrm{EB Garamond}
%\setmathfont(Digits,Latin)[Numbers={Lining,Proportional}]{EB Garamond}

\definecolor{linenocolor}{gray}{0.5}
\renewcommand\thelinenumber{\color{linenocolor}\arabic{linenumber}}

\usepackage{fix-cm}

\setcounter{totalnumber}{10}

\newcommand{\mr}{\mathrm}
\usepackage{hyperref}
\usepackage[normalem]{ulem}
\useunder{\uline}{\ul}{}
\makeatletter
\newenvironment{startquote}[2][2em]
  {\setlength{\@tempdima}{#1}%
   \def\chapquote@author{#2}%
   \parshape 1 \@tempdima \dimexpr\textwidth-2\@tempdima\relax%
   \itshape}
  {\par\normalfont\hfill--\ \chapquote@author\hspace*{\@tempdima}\par\bigskip}
\makeatother
%\renewcommand{\baselinestretch}{1.5}

\title[STRATEGIC SOCIAL LEARNING]
{\large STRATEGIC SOCIAL LEARNING AND THE POPULATION DYNAMICS OF HUMAN BEHAVIOR:
THE GAME OF GO}
\author{Bret A. Beheim\(^1\), Calvin Thigpen\(^2\), Richard McElreath\(^3\)}
\address{\(^1\)Department of Anthropology, University of New Mexico.
\(^2\)Institute for Transportation Studies, University of California,
Davis. \(^3\) Department of Anthropology, University of California,
Davis}
\email{beheim@gmail.com}
\date{March 28, 2014}

\begin{document}
%\begin{bibunit}[naturemag]

%{\abstract \small . }

%\renewcommand*\rmdefault{Minion}

\linenumbers
\modulolinenumbers[5]

\maketitle

%ABSTRACT
%\begin{abstract}\normalsize
\noindent\textsc{Abstract.}
Human culture is widely believed to undergo evolution, via mechanisms
rooted in the nature of human cognition. A number of theories predict
the kinds of human learning strategies, as well as the population
dynamics that result from their action. There is little work, however,
that quantitatively examines the evidence for these strategies and
resulting cultural evolution within human populations. One of the
obstacles is the lack of individual-level data with which to link
transmission events to larger cultural dynamics. Here, we address this
problem with a rich quantitative database from the East Asian board game
known as Go. We draw from a large archive of Go games spanning the last
six decades of professional play, and find evidence that the
evolutionary dynamics of particular cultural variants are driven by a
mix of individual and social learning processes. Particular players vary
dramatically in their sensitivity to population knowledge, which also
varies by age and nationality. The dynamic patterns of opening Go moves
are consistent with an ancient, ongoing arms race within the game
itself.
%\end{abstract}

\vspace{12pt}

\vspace{24pt}
% \newpage

\section{Introduction}\label{introduction}

One of the greatest contributions of evolutionary theory is the ability
to answer ``why'' questions about living things, explaining the
existence of their intricate and beautiful designs as adaptations via
natural selection. This success extends to study of our own species; we
now know a great deal about our behavior, psychology and, more
indirectly, our ancestry, as a consequence of evolutionary theory
\citep{Barkow1992adaptedmind, LalandBrown2002sense}. Many pathological
or idiosyncratic human social phenomena, for example, can be understood
as a mismatch between the environment our minds and bodies are
``expecting'' (subsistence foraging and horticulture) and the one many
of us live in (industrial food markets and sedentary lifestyles).

It is less clear, though, how much evolutionary theory helps us
understand the dynamics of human history, in which technology,
institutions and ideas change very rapidly. Work in the last three
decades has approached this problem by treating culture as a system of
inheritance, akin to but distinct from genetic transmission
\citep{Durham1992coevolution, Richerson2005}. Mathematical models have
shown that such systems will exhibit Darwinian adaptive dynamics even if
the traits under study are not discrete, particulate replicators like
genes, and even if they are not created through random mutation
\citep{boyd1985culture, Henrich2002}.

As with genetic evolution, the driving processes are organized within a
set of general categories we call ``forces'', but unlike natural
selection, genetic drift, meiotic drive, etc., the forces of cultural
evolution are largely based on the details of human psychology. What
humans find easy or appealing to learn, what cues and associations they
find salient, and how readily certain cultural traits encourage their
own social transmission. As a result, much research in cultural
evolution has concentrated on human social and individual learning
\citep{Mesoudi2006}, and has shown that simple learning heuristics, such
as ``when in doubt, copy what the majority is doing'' are plausible
explanations for how humans efficiently identify fitness-enhancing
behaviors \citep{McElreathetalSuccessbias2008}.

These learning processes, in aggregate, change the population
distribution of cultural variants in a recursive fashion, and in turn
alter the environment for future social learning. Thus, by applying what
Ernst Mayr called ``population thinking'' to human culture, it is
possible to develop mechanistic theories of how socially-transmitted
technologies, beliefs and behaviors spread and decline over time
\citep{shennan2009pattern}.

Though these learning models have been supported by laboratory and field
experiments in which participants are given controlled access to
potentially valuable social information
\citep{baum2004cultural, caldwell2008studying, effersontakezawa2007, Mesoudi2011a},
studying cultural evolution outside controlled settings is a much harder
task. \citet{michel2011quantitative} use Google book archives to trace
word use frequencies over the last several centuries but, lacking
information about who is using those words and under what circumstances,
they are unable to explain these trends beyond phenomenological
descriptions and qualitative associations to major historical events.
Ideally, if we had individual-level phenotypic measures for an entire
population over many years, we could identify the driving forces of
cultural change, as has been accomplished in non-human evolutionary
systems (e.g. \citet{grant2002unpredictable, ozgul2009dynamics}).

In what follows, we present an analysis of such a high-resolution
(\(n=31,137\)) archive of cultural patterns: records from the East Asian
game of Go. Over many decades of professional play, we can see
particular strategies and counterstrategies be innovated, copied,
decline, and vanish altogether. Our results provide strong evidence
that, on the average, these trends are the result of strategic social
learning decisions by professional Go players as they adopt and abandon
variant openings. This learning model, in aggregate, is the basis for an
apparent evolutionary arms race that is largely responsible for the
oscillations in Go openings among professional players. In this way, we
are able to link individual-level strategic learning to aggregate
population dynamics.

\section{Historical Dynamics in the Game of
Go}\label{historical-dynamics-in-the-game-of-go}

The East Asian board game known in the West as Go is, by players, one of
the most popular in the world, and certainly one of the oldest living
games. Originating in China between two and three thousand years ago
(where it is called \textit{weiqi}), it has since diffused to Korea
(there the game is called \textit{baduk}) and Japan (\textit{go} or
\textit{igo}). Through the Japanese, it became widely known in the West
in the early 20th century, and most American and European Go terminology
is derived from Japanese. With many centuries of paid professional
leagues and an abundance of game records (see Fig. 1), Go is an ideal
``model system'' for the quantitative study of human behavioral change
\citep{gobet2004moves}.

\begin{figure}[t]
  \begin{center}
    \includegraphics[scale=1.3]{./figKifu.pdf}
    \caption{A complete record, with SGF coordinates, for a 1978 game between Japanese professionals Takemiya Masaki and Kudo Norio. Black (Takemiya) begins with a Fourfour at coordinate \textit{pd} (top right), with White (Kudo) responding at \textit{dd} with another Fourfour. White won by resignation after 232 moves. This game is one of the first successful uses of the \textit{dd} response to the Move 1 Fourfour, and \textit{dd} has subsequently become the dominant response among professionals (see Fig. 3).}
    \label{FigKifu}
  \end{center}
\end{figure}

The rules of the game are easily explained: two players (Black and
White) sit on opposite sides of a large 19x19 coordinate grid. Starting
with Black, each player sequentially places a single stone on an empty
grid intersection of their choice. At the end of the game, the winner is
the player who controls the majority of the board (that is, whose stones
surround or occupy the most grid intersections). If the eventual outcome
is obvious, one player may resign mid-game; otherwise, once all possible
moves have been made, the players count grid intersections to see who is
the winner.

Beyond this basic outline of the game, there are two key rules
regulating play. First, stones or groups of stones of one player that
become completely surrounded on the grid by stones of the other are
removed from the board (this is the only time stones are moved, once
placed). Second, to prevent the possibility of the game entering an
infinite loop, the \emph{ko} rule restricts play so that the same board
state cannot appear twice in one game.

Though the rules above are remarkably simple, the game is fantastically
complex, and it takes years of dedicated study to become a competitive
player. Much of one's skill in Go comes from the ability to make
judgments about the long-term influence of particular moves and
formations on the board, a task that currently bedevils attempts at
developing strong computer opponents \citep{rimmel2010current}.

The first few moves of game of Go, which allow the most possibilites,
are actually the most predictable. The opening period of the game is
often characterized by highly normative, canalized play, where players
often follow one of a set of memorized move sequences called
\emph{joseki} (or \emph{jungsuk} in Korea), thought to grant neither
player a major advantage. Once a particular \emph{joseki} pattern has
started, professional players will draw on an encyclopedic knowledge of
explored routes in deciding their next move.

Despite the conservatism of opening play in professional games, novel
variations appear regularly and, if widely copied, become new
\emph{joseki} in the canon. Go historian Peter Shotwell writes of the
evolution of \emph{joseki}: ``Some are hundreds (or thousands) of years
old; others are born, die, and then are reborn again as the style of
play changes; still others are yet to be born and patiently wait their
turn to be discovered in some flash of ingenuity or desperation.''
\citep{shotwell2011go}.

\begin{figure}[t]
  \begin{center}
    \includegraphics[scale=1.2]{./figM1FreqsWins.pdf}
    \caption{The ebb and flow of the three major first move variants, with proportion of games won. The Fourfour's rise and fall, here broken into four eras, closely tracks with its relative performance (\textit{top}) since 1967 ($r=0.45$). Similarly, the Threethree's brief efflorescence and extinction closely match its relative win rate. Because of its initial dominance, all win rates are expressed relative to the Threefour's, and the dotted gray line indicates the location of the 50\% (break-even) point.}
    \label{FigM1FreqsWins}
  \end{center}
\end{figure}

\begin{figure}[t]
  \begin{center}
    \includegraphics[scale=1.2]{./figM2FreqsWins.pdf}
    \caption{Frequency and performance for four major M2 responses to the Fourfour. Performance for each is calculated as the win rate subtracting the average win rate of alternatives. As at M1, a variants's frequency closely tracks its relative performance. To compare with Fig. 2, we define four eras (I to IV) corresponding to the Fourfour's rise and fall.}
    \label{FigM2FreqsWins}
  \end{center}
\end{figure}

\clearpage

As shown in Fig. 2, we can see such dynamics of the very first move
(Move 1, or M1) over the latter half of 20th century professional play.
Once board symmetry is taken into account, there are only two common
locations Black will play, the Fourfour and the Threefour, each named
after their coordinate locations on the corners of the board.

After a brief enthusiasm for the Threethree in the 1960s, the Fourfour
rose to dominance, peaking in 1977 (appearing in 78\% of games played)
and again in 1996 (82\% of games played). The primary goal of this
analysis is to understand why this pattern exists, and what processes
are driving evolution at M1.

We can trace the rapid diffusion of the Fourfour during this time back
to the success of a single innovator, Japanese professional Takemiya
Masaki. Beginning his pro career at the age of 14 in the 1960s, Takemiya
has used the Fourfour almost exclusively; of his 520 games as Black in
the database, only six do not start with the Fourfour. Takemiya's novel
use of this variant in 1968 marks the beginning of its spread within the
broader Go community, and his personal performance for the next five
years (64\%) also exceeded the average Black win rate (61.5\%).

Subsequent to Takemiya's arrival, the success of the Fourfour appears to
have prompted an innovation race in White's Move 2 response. Before
1970, White's response to the Fourfour was almost always a Threefour on
an adjacent corner. If we imagine the M1 Fourfour is played in the upper
right quadrant (SGF coordinate \emph{pd}), nearly 80\% of the time the
M2 response in the late 1960s was coordinate \emph{dc} (Fig. 3). By
1977, at the Fourfour's first peak, the \emph{dc} response declined to
about 30\% of games. In the interim, M2 variants at \emph{dp}, \emph{cq}
and \emph{dd} all became established in rapid sequence. As measured by
the proportion of games White won, \emph{dp} and \emph{cq} both
consistently outperformed \emph{dc} in the 1970s, blunting the
Fourfour's effectiveness and anticipating its decline in use after 1977.

\subsection{Strategic Social Learning}\label{strategic-social-learning}

From the above, it appears that the waxing and waning success of the
Fourfour stems from an ongoing crucible of innovation and
counter-innovation. As it rapidly proliferated in the professional
community in the 1970s, the Fourfour enjoyed an anomalously large win
rate. Conversely, the Fourfour's decline after 1977 occurred when its
win rate was consistently below that of its rival variant, the M1
Threefour.

With so many possible opening moves, the few that are observed over the
decades of play suggest an important role for cultural transmission from
player to player through the generations. But is social learning
responsible for the oscillations in move frequencies observed in Figs. 2
\& 3?

As strategic actors in a high-stakes competitive environment,
professional Go players are under immense pressure to choose effective
moves that help defeat their opponents. We should expect that players
update their valuation of particular sequences through trial and error
in their own games, so that past performance predicts current behavior.

At the same time, the complexity of the game, dynamic strategy space,
and high cost of individual experimentation make objective conclusions
about the current value of particular moves very difficult to establish.
As such, it is also plausible that players may strategically tap into
the knowledge and experience contained in the games of other players,
records of which are widely available (with many of the games taking
place at the same Go institutes in China, Japan and South Korea).

Theoretical models and experiments have focused on two important forms
of such strategic social learning: frequency-dependent and
success-biased transmission
\citep{boyd1985culture, McElreathetalSuccessbias2008}.
Frequency-dependence is one of the simplest kinds of social learning, in
which learners use the popularity of different behaviors to decide which
to adopt. It can arise through an unbiased or biased sampling scheme,
and by copying the behavior of others, individuals can quickly and
cheaply acquire useful information that would otherwise require costly
individual learning.

\begin{figure}[t]
  \begin{center}
    \includegraphics[scale=1.2]{./FigPr44xPop_COL.pdf}
    \caption{Predicted probability of using the Fourfour by an average player at different values of population use, with means (dotted lines) and 95\% HPDI (solid regions). Per Table 1, the relative win rate of the Fourfour within the population regulates the effect of population use, -10\% versus the Threefour (orange), equal to the Threefour (light blue) and +10\% (dark blue). Outcome data is also shown via scatterplot.}
    \label{FigPr44}
  \end{center}
\end{figure}

Success-biased social learning exploits the correlation between
behaviors and outcomes, so that a move that is associated with greater
success among others may be favored simply for that reason, regardless
of who is using it, how popular it is, or whether the learner really
understands why it works.

Using statistical modeling techniques, we can evaluate how consistent
real players are with these idealized models. Each of these three
concepts, individual updating, frequency-dependent and success-biased
social learning, were formalized as multilevel logistic regression
models for the probability any particular game will start with a
Fourfour.

\section{Methods}\label{methods}

We model the decision to use the Fourfour variant at M1 via a binomial
model with a logistic link function of the probability of use, fitted by
Hamiltonian MCMC in the STAN software package \citep{stan2013stan} and
the R programming language. For each game \(i\), we model whether Move 1
was a Fourfour (\(y_i=1\)) or not (\(y_i=0\)) via a multilevel logistic
regression framework. We used two classes of predictors, ``Personal
Knowledge'' and ``Population Knowledge''. Personal variables include a
player's recent (last two years) use of the Fourfour for Move 1, and how
often they won games with it, relative to how often they won games
without it. This model corresponds to the hypothesis that players update
their views of the move, based on how familiar it is to them and how
many games they have won using it. ``Population Knowledge'' models use
various predictors constructed from the recent population use of the
Fourfour, and the population win rate of the Fourfour relative to its
alternatives. These models attempt to measure frequency-dependent and
success-biased social learning described above.

Predictor variables for each game used a retrospective two-year window
from the date the game took place, taking the simple average of values
for relevant games within that period. Each predictor representing a use
frequency was centered on 0.5, a natural point of interpretation for
whether a behavior is in the majority. Predictors for win rate, in
contrast, were centered on the average win rate of alternative variants
at that game move.

\begin{table}[t]
  \centering
    \begin{tabular}{lrr}
    \input{"./table1.txt"}
    \end{tabular}%
  \caption{Logistic regression estimates for Pr(Use Fourfour in game $i$)}
\end{table}

\section{Results}\label{results}

\begin{figure}[t]
  \begin{center}
    \includegraphics[scale=0.8]{./FigPlayerAgeBetaGamma_COL.pdf}
    \caption{(\textit{left}) Six player-specific joint posterior densities of the importance of individual ($\beta_j$) and social ($\gamma_j$) information on predicting move outcome, along with the main effect density (black region). Players who lie above the line of equality (solid gray) are more predicted by recent population trends than individual experience, while the reverse is true for players below it. (\textit{right}) Estimates for the importance of social knowledge for each year of age $k$, showing means and 95\% HPDI intervals. Younger players tend to have values of $\gamma_k$ above the main effect mean $\gamma$ (dotted), but this declines as they age.}
  \label{FigPlayerAgeBetaGamma}
  \end{center}
\end{figure}

With all predictors centered on means, the intercept estimates that an
average Go player will use the Fourfour about 53.2\% of the time during
the study period (i.e.~logit(.532)=0.13). A player's recent use of a
particular opening move strongly predicts their future use; for an
otherwise-average player who used the Fourfour about 60\% of their
recent games, the probability they will use the Fourfour rises to 58.7\%
(OR: 9.39). There is a smaller (OR: 2.77), but unambiguous, positive
performance effect from past success with the Fourfour, adding about two
percentage points to the probability of use for a player who has won
10\% more games using the Fourfour than his overall win rate. These two
personal effects also interact, so that past success enhances the
predictive power of past use (or vis versa).

Most important to the question of social learning and cultural
evolution, though, is whether there is predictive value in the recent
behavior of \emph{other} players, controlling for individual behavior.
Indeed, when the Fourfour appears in 60\% of \emph{all} recent games,
the expected probability of use by an average player rises from 53.2\%
to 59.0\% (OR: 10.475). Two important interaction effects regulate the
frequency-dependence of a player's behavior; if the Fourfour appears in
60\% of recent professional games, \emph{and} its population performance
is ten percentage points higher than that of its rival, the Threefour,
the expected probability a player will use it rises from 59.0\% to
61.4\% (Fig. 4).

Because the coefficient for personal use (which we call \(\beta\)) and
that for population use (\(\gamma\)) have the same units, comparing
their magnitudes tells us the relative predictive value of personal and
population knowledge. For an average player, social knowledge is 1.05
times more important than personal experience in predicting the outcome,
but this belies the enormous variation from player to player. Partially
pooling games of each Black player, we see a varying effects standard
deviation of 4.00 for the distribution of player-specific \(\beta_j\).
For population knowledge, the standard deviation across player
intercepts is even larger, 9.77. Plotting the joint posterior densities
of \(\beta_j\) and \(\gamma_j\) for each player \(j\) (Fig. 5, left)
reveals that some professionals are substantially better predicted by
social knowledge than by their own experiences, while other players are
almost completely insensitive to population trends (foremost among them,
Takemiya Masaki).

Including player age as a varying effect regulating magnitude of
\(\beta\) and \(\gamma\) also shows learning heterogeneity over the life
course (Fig. 5, right). The games of the youngest professionals, those
between the ages of 14 and 18, are least predicted by personal
experience and most by population trends. After age 20, the balance
steadily shifts towards personal experience, peaking at age 27 well
above the average \(\beta\) of 2.24. Players older than 35 appear to
return to average levels of social learning; with a few deviations, no
trend is apparent after this age.

\section{Discussion}\label{discussion}

Though only a preliminary treatment of a very rich dataset, the above
results demonstrate how a socially-transmitted cultural pattern of
behavior - the first move variants in the game of Go - undergo
evolutionary change. In deciding which opening move to use, it appears
Go professionals attend to both their own past experiences and those of
other players, placing substantial weight on the recent popularity and
performance of different variants.

As each player's decision and subsequent game outcome adds new
information to the body of population knowledge, this process proceeds
recursively, driving the observed cyclic trends in move frequencies. We
can also see how the breakaway success of the Fourfour variant at M1 was
followed by a period of innovation and diversity in more effective M2
responses, consistent with an ongoing evolutionary arms race conducted
through strategic social learning.

According to player-specific posterior estimates in Fig. 5, the players
who most strongly track population trends are South Korean professionals
Lee Sedol and Cho Namchul, while Chinese player Peng Quan and Japanese
player Takemiya Masaki lie at the other extreme. This segregation of
nationalities may not be a coincidence; as Table 2 shows, the average of
the player-specific estimates for individual and population knowledge
differs across country, with South Korean players tending toward
greater-than-average sensitivity to social information (positive mean
\(\gamma_j\)). The average Japanese player, in contrast, is better
predicted by their individual knowledge (positive mean \(\beta_j\)),
with the 75 Chinese players nearest to the population intercepts
\(\beta\) and \(\gamma\), on the average. It is important to emphasize
that these distributions show substantial overlap, but this result
raises the tantalizing possibility that normative or institutional
differences across East Asia may encourage different forms of social
learning within each national league.

\begin{table}[t]
  \centering
    \begin{tabular}{lrrrrr}
    \input{"./table2.txt"}
    \end{tabular}%
    \caption{National averages of player-specific estimates of reliance on individual ($\beta_j$) and social ($\gamma_j$) information.}
\end{table}

Although we can do much to reconstruct and explore the historical
record, several important questions remain outstanding. First, we do not
know the nature of the observed success bias. We can see that relative
performance of the Fourfour versus the Threefour consistently improves
our predictions of which one a particular player will use. This seems to
indicate players focus on how well a move does, but it is equally
possible, however, that successful players rather than successful
variants are what are emulated.

This is not captured by crude measures like population use frequency,
and is indicative of a larger problem: establishing exactly who players
are learning from. The statistical models above treat the general
population as an equally-connected, homogenous pool. Real social
influence in this situation undoubtedly involves a variety of complex
social relationships between unique personalities. The nature of these
influences may be better captured by network-based approaches that allow
for heterogeneous interconnections between players. It may be possible
to use information about players' nationalities or recent matches to
develop measures of social proximity, allowing the detection of
theoretical effects like conformist learning or prestige.

Another problem is the nature of innovation in Fourfour play. Even if
Takemiya's innovative opening style (widely known as ``Cosmic Go'') was
the catalyst for the Fourfour's rise in the 1970s, as the evidence above
seems to indicate, we do not know if other adopters were using it in the
same way. It could be, for example, that Takemiya's early successes in
1968-1970 encouraged other professionals to revisit the Fourfour and
develop their own innovative applications, rather than emulate his.

Finally, even if players really are using a combination of individual
and population knowledge, we should consider the possibility that this
offers no real strategic advantage. A well-understood problem with
success bias is that it aims to exploit an association between use and
performance, but can lead to the proliferation of neutral or even
deleterious practices if that association is not causal. Rather than
produce adaptation, social learning may simply be diffusing the
(arbitrary) behaviors of a high-performing elite.

There are many more possible avenues of exploration in this very rich
dataset. Venturing beyond Move 1, we can study the diffusion of
particular joseki patterns, the importance of particular high-status
individuals or, indirectly, the diffusion of general strategies within
professional play. The scope of study can also be expanded; large
archives of professional games go back at least several hundred years,
and tens of thousands of amateur games are available online, to say
nothing of similar archives available for other games like chess.

In any case, the statistical and theoretical tools employed above can be
used to expand our understanding of cultural dynamics. The above results
hopefully demonstrate the value of adopting a quantitative, Darwinian
point of view towards cultural change. As in evolutionary ecology,
knowledge gained from high-resolution datasets can carry over to human
historical dynamics for which little data remains and variants are
harder to define.

Indeed, we feel that the theoretical and empirical evidence is so
overwhelming that the key question is no longer whether culture evolves,
but rather which evolutionary processes operating in day-to-day life
brought about the particular patterns of cultural diversity we see in
human history.

\bibliographystyle{apalike}\bibliography{citations}

\newpage

\section{Appendix}\label{appendix}

\subsection{The GoGoD Database}\label{the-gogod-database}

Games were pulled from the 2009 Games of Go on Disk (GoGoD) library,
published by T Mark Hall and John Fairbairn. Data is drawn from all
database games within the years 1954 to 2009, but because predictors use
a two-year retrospective window from the day and month the game was
played, the first games analyzed took place in 1956. From this year
range, we excluded all games that took place with handicap stones (in
which White plays first, with Black stones already on the board). We
also excluded exhibition games with unusual rules, games with known
corruptions described in the comments, and games in which one of the
players was an amateur by rank, rather than a paid professional (the one
exception to this rule was Kikuchi Yasuro, b. 1929, who played evenly
with top professionals).

Since we are using player-specific predictors and grouping games by
player in our varying effects, we excluded games whose Black players did
not have at least 50 games on record between 1956 and 2009. After these
exclusions, a total of 31,137 games remained within the database, with
207 unique players taking Black. In addition to the data contained
within the game record and the metadata, we also collected player
nationality and date of birth was drawn from a variety of sources,
chiefly player biography pages from the Go wiki ``Sensei's Library''
(http://senseis.xmp.net/). Although the GoGoD database is itself
commercially licensed by Fairbairn and Hall, the regression table used
here is available by request to the corresponding author.

\subsection{Full Regression Model}\label{full-regression-model}

Our basic question is whether or not a player will use the Fourfour
opening move in any particular game, modelled as a binomial random
variable with a logistic probability model. This regression model has
three important features: terms relating to personal information, terms
relating to population information, and stochastically modelled
player-specific intercepts and slopes. Specifically, if the probability
of using the Fourfour in game \(i\) of Black player \(j\) who is aged
\(k\) is \(p_{ijk}\), then the log-odds that the Fourfour opening move
is used is modelled by

\begin{equation} \label{eq:model}
  \log\bigg(\frac{p_{ijk}}{1-p_{ijk}}\bigg) = \alpha + \alpha_j + \beta_{ijk}x_i + \gamma_{ijk}y_i + \kappa q_i + \eta r_i + \textrm{controls}.
\end{equation}

Predictor \(x_i\) is the proportion of Black player \(j\)'s games that
use the M1 Fourfour recently (within the last two years), \(y_i\) is the
frequency of the M1 Fourfour among recent games within the general
population, \(q_i\) is that player's recent win rate specifically among
their Fourfour games, and \(r_i\) is the win rate among all recent
Fourfour games in the population. Of particular interest are parameters
\(\beta_{ijk}\) and \(\gamma_{ijk}\), which represent the importance of
personal and population use frequencies in the model. These two
parameters are themselves modelled by \[
  \beta_{ijk} = \beta + \beta_j + \beta_k + \beta_{S} s_{i} + \beta_{Q} q_{i}
\] \[
  \gamma_{ijk} = \gamma + \gamma_j + \gamma_k + \gamma_{S} s_{i} + \gamma_{R} r_{i}
\]

In addition to interactions for \(q_i\) and \(r_i\), which allow the
Fourfour's recent win rate for the player and the population to regulate
the importance of its frequency, predictor \(s_i\) is the Black player's
overall win rate among games played in the last two years. Within the
model, intercept \(\alpha_j\) and slopes \(\beta_j\) and \(\gamma_j\)
for player \(j\) treated as varying effects following a joint normal
distribution with mean \textbf{0}, as are those of players of age \(k\).

The controls for each game \(i\) are made up of six predictors: White's
win rate using the M1 Fourfour in past games (where they played as
Black), White's loss rate \textit{against} the M1 Fourfour, White's
frequency of use of the Fourfour, the number of past games Black has
seen the Fourfour used, whether or not Black lost the last time someone
used an M1 Fourfour on him, and the handicap of game \(i\).

This model was first fit using the lme4 package in the R language and by
Hamiltonian Monte Carlo via the STAN software package called by the
glmer2stan command (McElreath 2013), running 5,000 iterations with a
1,000 iteration burn-in. Convergence of chains was determined by the
R-hat Gelman \& Rubin diagnostic \citep{stan2013stan} and by visual
inspection of the chains.

\end{document}